---
title: 'Upgrading to R #1: Excitement >> Skills'
author: Alex K Gold
date: '2019-01-20'
slug: upgrading-to-r-1-excitement-skills
categories:
  - conferences
  - R
  - RStudio
tags: []
---
*This is the first in a series of posts sharing some insights on upgrading to R. See the [index post](/2019/01/19/upgrading-to-r-at-rstudio-conf-2019/) for others.*
```{r setup, include=FALSE}
library(tidyverse)
library(gganimate)
source(here::here("content", "post", "rsconf2019", "helper_funcs.R"))
knitr::opts_chunk$set(echo = FALSE, fig.width = 4)
options(gganimate.start_pause = 15)
```

Walking on to an R-less team on the first day is a daunting experience, made all the more so by not knowing whether you've got the right people. In my experience, there are a few factors that can help drive the team from here (a sadly R-less land) to the magical place where free range R code runs all day, the data is always clean, and syntax errors are nonexistent. 

Here's my estimate of the effect sizes of some important factors:
```{r skills, echo = FALSE, messge = FALSE, warning =FALSE}
skills = c("Tidyverse-First Orientation",
           "Relentless R",
           "Positive Learning Environment", 
           "Team's Enthusiasm for R",
           "Starting Level of R Knowledge")

options(gganimate.nframes = 100, 
        gganimate.end_pause = 30)
labs <- skills %>%
  stringr::str_wrap(20)

g <- tibble(x = seq(length(labs)),
       pe = c(0.2, 0.33, 0.4, 0.7, 0.1),
       ymin = c(0.1, 0.15, 0.2, 0.4, -0.2),
       ymax = c(0.4, 0.4, 0.48, 0.9, 0.2),
       lab_y = c(0.8, 0.8, 0.8, 0.2, 0.8),
       lab = labs,
       time = x) %>%
  ggplot(aes(x = -x, y = pe, group = lab)) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax)) +
  geom_text(aes(y = lab_y, label = lab), family = text_fam) +
  theme_rs() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  geom_hline(yintercept = 0, linetype = 2, color = "lightgrey") +
  ylab("Effect on Team's Productivity in R") +
  coord_flip(xlim = -c(0.5, 5.5)) +
  transition_reveal(time) +
  enter_fade() +
  ease_aes('cubic-in-out') +
  labs(caption = "Completely fake data.")
animate(g, width = 450, height = 450)
```

## Wired: Team's enthusiasm for learning R
Tired: How much R they already know

This is pretty straightforward. By far the most important factor that will determine whether your teammates successfully join you in transitioning to R is whether they're excited about it, and the least important is how much they already know. Learning R just really isn't that hard (and I'll share some tips to make it easier for your team). People who are excited to learn will do it, and those who aren't won't. It's that easy. 

Don't know who's excited and who's not? Ask them! And don't just ask whether they're excited to learn R. Ask them whether they're excited to work in R as a team, to adopt the tools and workflows that will lead to success *as a team*. Transitioning to R has a bunch of benefits, but reaping them all requires a lot more than just writing some code. It's a wholesale transition in terms of how the team thinks about work.

## Lesser Angels
Beyond enthusiasm, there are a couple of things you can do to make it easier, whether you're a team member or the leader.

* *Tidyverse-First Orientation*: If you've been doing R for a while, you probably know this, but the Tidyverse is AWESOME. It takes the wonderful benefits of R and adds an opinionated take on what code should look like. If you're just learning R, start with the Tidyverse. You'll be glad you did because it's [magical](https://docs.google.com/presentation/d/1iUz3A_xHIzeFIzOQ_JQLdGPbC82qcDEDSqvjIBiP1a8/edit). [^1]
* *Relentless R*: What you're doing isn't coding, it's organizational change. Writing code is easy. Changing hearts, minds, and workflows is _hard_. There are people who make whole careers out of this stuff. Don't get down if it's slow going. I'm skeptical it's possible to really transition a team to R in less than 12 months.
* *Positive Learning Environment*: You're trying to make room for people to learn. That means it's important to make work a space where they won't feel stupid for asking questions. Hopefully this is already true for your workplace...but it's especially important if you're trying to get your team to do something new and hard.

[^1]: If you know R, and you're trying to get other people onboard, just pass along the good word.

